ive cover csv---
title: "Lupine manuscript"
author: "Becca Evans"
date: "2023-04-19"
output: html_document
---

Vegetation cover
```{r}
library(lattice)
library(ggplot2)
library(cowplot)
library(dplyr)
library(lme4)
library(lmerTest)
library(emmeans) #allows post hoc tests for lmer
```

#data
#veg survey vegetative cover.csv

#models
###leafminer damage####
```{r}
hist(lupdata.byplot$lfmnr.pct)

#sqrt transform improves residual behavior and results in lower p-value
mod1 <- lmer(sqrt(lfmnr.pct) ~ treatment + (1|block), data=lupdata.byplot)
anova(mod1)
summary(mod1)
plot(mod1)
qqnorm(resid(mod1)) #excellent
lsmeans(mod1, list(pairwise ~ treatment), adjust = "tukey")

```

#####lupine percent cover#####
```{r}
hist((lupdata.byplot$lup.pct))
mod2 <- lmer((lup.pct) ~ treatment + (1|block), data=lupdata.byplot)
anova(mod2)
summary(mod2)
lsmeans(mod2, list(pairwise ~ treatment), adjust = "tukey")
lsmeans::lsmeans(mod2, "treatment")
plot(mod2) 
qqnorm(resid(mod2))  
qqline(resid(mod2))
```



###vascular cover###
```{r}
hist((lupdata.byplot$vasc.pct))
mod3 <- lmer((vasc.pct) ~ treatment + (1|block), data=lupdata.byplot)
anova(mod3)
summary(mod3)
lsmeans(mod3, list(pairwise ~ treatment), adjust = "tukey")
lsmeans::lsmeans(mod3, "treatment")
plot(mod3)

```


#Moss
```{r}
hist((lupdatat.byplot$moss.pct))
mod4 <- lmer((moss.pct) ~ treatment + (1|block), data=lupdata.byplot)
anova(mod4)
summary(mod4)
plot(mod4)

```

#bare
```{r}
hist((lupdata.byplot$bare.pct))
mod5 <- lmer((bare.pct) ~ treatment + (1|block), data=lupdata.byplot)
anova(mod5)
summary(mod5)
lsmeans::lsmeans(mod5, "treatment")
plot(mod5)

```


###soil C,N, d13c, d15n, microbial C

#data=soil.csv

#models
 
Full model response~treat*month*location*depth*loc+(1|block)
use step() to select best models presented here 

soil C
```{r}
soil_c <- lmer(log(c)~treat+month+loc+depth+month+treat:loc+treat:depth+(1|block),data=data)
anova(soil_c)
summary(soil_c)
plot(soil_c) 
qqnorm(resid(soil_c))  
```

post hoc soil c loc
```{r}
emmeans(soil_c, pairwise ~ treat | loc, type="response")
```

post hoc soil c depth 
```{r}
emmeans(soil_c, pairwise ~ treat | depth, type="response")
```

soil n NO EFFECTS
```{r}
soil_n <- lmer(log(n)~treat+loc+depth+month+(1|block),data=data)
n <- anova(soil_n)
summary(soil_n)
plot(soil_n) 
qqnorm(resid(soil_n)

```

soil d13c
```{r}
soil_d13c <- lmer(d13c~treat+month+loc+depth+month+treat:loc+treat:depth+(1|block),data=data)
c <- anova(soil_d13c)
summary(soil_d13c)
plot(soil_d13c) 
qqnorm(resid(soil_d13c))  

```


post hoc soil d13c loc
```{r}
emmeans(soil_d13c, pairwise ~ loc, type="response")
```

soil d15n NO EFFECTS
```{r}
soil_d15n <- lmer(d15n~treat+month+loc+depth+month+treat:loc+treat:depth+(1|block),data=data)
a <- anova(soil_d15n)
summary(soil_d15n)
plot(soil_d15n) 
qqnorm(resid(soil_d15n))  

```


microbial C
```{r}
microbe_c <- lmer(sqrt(microbial.C+1)~treat+month+loc+depth+treat:loc+treat:depth+treat:month+(1|block),data=data)
m <- anova(microbe_c)
summary(microbe_c)
plot(microbe_c) 
qqnorm(resid(microbe_c)) 

```


post hoc microbial c treatment
```{r}
emmeans(microbe_c, pairwise ~ treat, type="response")
```

post hoc microbial c loc*treat interaction
```{r}
emmeans(microbe_c, pairwise ~ loc|treat, type="response")
```

post hoc microbial c loc interaction
```{r}
emmeans(microbe_c, pairwise ~ loc, type="response")

```

post hoc microbial c date interaction
```{r}
emmeans(microbe_c, pairwise ~ month, type="response")
```


##look at just interspaces treatment effects
```{r}
int <- data %>% filter(loc=="Interspace")
microbe_c_int <- lmer(sqrt(microbial.C+1)~treat+month+depth+treat:depth+treat:month+(1|block),data=int)
m <- anova(microbe_c_int)
write.csv(m, "m.csv")
summary(microbe_c_int)
plot(microbe_c_int) 
qqnorm(resid(microbe_c_int)) 

```

post hoc microbial c treat 
```{r}
emmeans(microbe_c_int, pairwise ~ treat, type="response")
```

##look at justunder lupine treatment effects
```{r}
lup <- data %>% filter(loc=="Under Lupine")
microbe_c_lup <- lmer(sqrt(microbial.C+1)~treat+month+depth+treat:depth+treat:month+(1|block),data=lup)
m <- anova(microbe_c_lup)
write.csv(m, "m.csv")
summary(microbe_c_lup)
plot(microbe_c_lup) 
qqnorm(resid(microbe_c_lup)) 

```

post hoc microbial c treat 
```{r}
emmeans(microbe_c_lup, pairwise ~ treat, type="response")
```

###############
####microbial community composition 

#data
micro_july <- read_csv("diversity_july1.csv") %>% 
  column_to_rownames("site")

micro_aug <- read_csv("diversity_aug1.csv") %>% 
  column_to_rownames("site")

july_env_full <- read.csv("bulk_soil_BE 24Jun2018_july.csv")
aug_env_full <- read.csv("bulk_soil_BE 24Jun2018_aug.csv")

###richness

sppr <- specnumber(micro_july)
sppr

sppr_mod <- lmer(sppr~treat+(1|block), data=july_env_full)
summary(sppr_mod )
anova(sppr_mod )

#plot
sppr_df <- sppr %>% 
  enframe() %>% 
  full_join(july_env_full, by = c("name" = "site"))

pal <- c("lightsalmon1", "gold1", "palegreen4")

plot_sppr <- ggplot(sppr_df, aes(x = treat, y = value, fill = treat)) +
  geom_boxplot() +
  scale_color_manual(values=pal, breaks=c("c","r","s"), labels=c("Control", "Lupine Rem.", "Herbiv.Exc"))+
  
  labs(x = "",
       y = "Number of PLFA")

     
plot_sppr

#shannon
shannondiv <- vegan::diversity(micro_july)
shannondiv

shan_div_mod <- lmer(shannondiv~treat+(1|block), data=july_env_full)
summary(shan_div_mod )
anova(shan_div_mod )

#plot

shan_df <- shannondiv %>% 
  enframe() %>% 
  full_join(july_env_full, by = c("name" = "site"))



pal <- c("lightsalmon1", "gold1", "palegreen4")

plot_shannon <- ggplot(shan_df, aes(x = treat, y = value, fill = treat)) +
  geom_boxplot() +
  scale_color_manual(values=pal, breaks=c("c","r","s"), labels=c("Control", "Lupine Rem.", "Herbiv.Exc"))+
  
  labs(x = "",
       y = "Shannon Diversity")




